\bf Visualising How Non-linear Dimension Reduction Warps Your Data

### 2024-01-14 paper.qmd
Add the prediction section where 2D embeddings are predicted.

#(knn: 10, init: random, \n MN-ratio: 0.9, FP-ratio: 2)
plot_list5 <- plot_PaCMAP_2D(PaCMAP_data) + #ggtitle("(e)") + 
  theme_linedraw() +
  theme(plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5), 
        axis.title.x = element_blank(), axis.title.y = element_blank(),
        axis.text.x = element_blank(), axis.ticks.x = element_blank(),
        axis.text.y = element_blank(), axis.ticks.y = element_blank(),
        panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
  annotate(geom = 'text', label = 'e', x = -Inf, y = Inf, hjust = -0.5, vjust = 1.5, size = 3)

@@

@@

In this context, the term "prediction values" refers to the 2D coordinates predicted for the NLDR technique. The approach involves employing the K-nearest neighbors (KNN) algorithm to identify the nearest hexagonal bin centroid in the 2D space. Subsequently, the coordinates of this centroid are assigned as the low-dimensional predicted values for the test data in 2D space. It is noteworthy that traditional NLDR methods, such as t-SNE, often lack a direct predict function, making our approach valuable for generating predicted values in the absence of such functionalities.

@@

>>>>

#(knn: 10, init: random, \n MN-ratio: 0.9, FP-ratio: 2)
plot_list5 <- plot_PaCMAP_2D(PaCMAP_data) + #ggtitle("(e)") + 
  theme_linedraw() +
  theme(plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5), 
        axis.title.x = element_blank(), axis.title.y = element_blank(),
        axis.text.x = element_blank(), axis.ticks.x = element_blank(),
        axis.text.y = element_blank(), axis.ticks.y = element_blank(),
        panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
  annotate(geom = 'text', label = 'e', x = Inf, y = Inf, hjust = 1.5, vjust = 1.5, size = 3)

@@

<!--model for S-curve with 5 methods-->

## Import data

tSNE_s_curve <- read_rds("data/s_curve/s_curve_tsne_27.rds")
UMAP_s_curve <- read_rds("data/s_curve/s_curve_umap.rds")
PHATE_s_curve <- read_rds("data/s_curve/s_curve_phate.rds")
TriMAP_s_curve <- read_rds("data/s_curve/s_curve_trimap.rds")
PaCMAP_s_curve <- read_rds("data/s_curve/s_curve_pacmap.rds")

num_bins_tsne_s_curve <- 8
shape_val_tsne_s_curve <- calculate_effective_shape_value(.data = tSNE_s_curve,
                                                          x = tSNE1, y = tSNE2) ## 1.259938
## To extract bin centroids
hexbin_data_object_tsne_s_curve <- extract_hexbin_centroids(nldr_df = tSNE_s_curve, num_bins = num_bins_tsne_s_curve, shape_val = shape_val_tsne_s_curve, x = tSNE1, y = tSNE2)

df_bin_centroids_tsne_s_curve <- hexbin_data_object_tsne_s_curve$hexdf_data

tSNE_data_with_hb_id_s_curve <- tSNE_s_curve |>
  dplyr::mutate(hb_id = hexbin_data_object_tsne_s_curve$hb_data@cID)

## To generate a data set with high-D and 2D training data
df_all_tsne_s_curve <- dplyr::bind_cols(training_data |> dplyr::select(-ID), tSNE_data_with_hb_id_s_curve)

## Averaged on high-D
df_bin_tsne_s_curve <- avg_highD_data(.data = df_all_tsne_s_curve)

## Triangulate bin centroids
tr1_object_tsne_s_curve <- triangulate_bin_centroids(df_bin_centroids_tsne_s_curve, x, y)
tr_from_to_df_tsne_s_curve <- generate_edge_info(triangular_object = tr1_object_tsne_s_curve)

# ggplot(df_bin_centroids_tsne_s_curve, aes(x = x, y = y)) +
#   geom_segment(data = tr_from_to_df_tsne_s_curve, aes(x = x_from, y = y_from, xend = x_to, yend = y_to)) +
#   geom_point(size = 2, colour = "#33a02c") +
#   coord_equal()

## Compute 2D distances
distance_tsne_s_curve <- cal_2D_dist(.data = tr_from_to_df_tsne_s_curve)

## To find the benchmark value
benchmark_tsne_s_curve <- find_benchmark_value(.data = distance_tsne_s_curve, distance_col = distance)

# colour_long_edges(.data = distance_tsne_s_curve, benchmark_value = benchmark_tsne_s_curve,
#                   triangular_object = tr1_object_tsne_s_curve, distance_col = distance)

trimesh_removed_tsne_s_curve <- remove_long_edges(.data = distance_tsne_s_curve, benchmark_value = benchmark_tsne_s_curve,
                                                  triangular_object = tr1_object_tsne_s_curve, distance_col = distance)

trimesh_removed_tsne_s_curve <- trimesh_removed_tsne_s_curve +
  # xlab(expression(C[x]^{(2)})) + ylab(expression(C[y]^{(2)})) +
  # theme(axis.text = element_text(size = 5),
  #       axis.title = element_text(size = 7)) +
  geom_point(colour = "#33a02c", size = 0.5) +
  theme_linedraw() +
  theme(plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5), 
        axis.title.x = element_blank(), axis.title.y = element_blank(),
        axis.text.x = element_blank(), axis.ticks.x = element_blank(),
        axis.text.y = element_blank(), axis.ticks.y = element_blank(),
        panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
  annotate(geom = 'text', label = 'a', x = -Inf, y = Inf, hjust = -0.5, vjust = 1.5, size = 3)

# tour_tsne_s_curve <- show_langevitour(df_all_tsne_s_curve, df_bin_tsne_s_curve, df_bin_centroids_tsne_s_curve, benchmark_value = benchmark_tsne_s_curve, distance = distance_tsne_s_curve, distance_col = distance)

## UMAP

num_bins_umap_s_curve <- 6
shape_val_umap_s_curve <- calculate_effective_shape_value(.data = UMAP_s_curve,
                                                          x = UMAP1, y = UMAP2) ## 1.259938
## To extract bin centroids
hexbin_data_object_umap_s_curve <- extract_hexbin_centroids(nldr_df = UMAP_s_curve, num_bins = num_bins_umap_s_curve, shape_val = shape_val_umap_s_curve, x = UMAP1, y = UMAP2)

df_bin_centroids_umap_s_curve <- hexbin_data_object_umap_s_curve$hexdf_data

UMAP_data_with_hb_id_s_curve <- UMAP_s_curve |>
  dplyr::mutate(hb_id = hexbin_data_object_umap_s_curve$hb_data@cID)

## To generate a data set with high-D and 2D training data
df_all_umap_s_curve <- dplyr::bind_cols(training_data |> dplyr::select(-ID), UMAP_data_with_hb_id_s_curve)

## Averaged on high-D
df_bin_umap_s_curve <- avg_highD_data(.data = df_all_umap_s_curve)

## Triangulate bin centroids
tr1_object_umap_s_curve <- triangulate_bin_centroids(df_bin_centroids_umap_s_curve, x, y)
tr_from_to_df_umap_s_curve <- generate_edge_info(triangular_object = tr1_object_umap_s_curve)

# ggplot(df_bin_centroids_umap_s_curve, aes(x = x, y = y)) +
#   geom_segment(data = tr_from_to_df_umap_s_curve, aes(x = x_from, y = y_from, xend = x_to, yend = y_to)) +
#   geom_point(size = 2, colour = "#33a02c") +
#   coord_equal()

## Compute 2D distances
distance_umap_s_curve <- cal_2D_dist(.data = tr_from_to_df_umap_s_curve)

## To find the benchmark value
benchmark_umap_s_curve <- find_benchmark_value(.data = distance_umap_s_curve, distance_col = distance)

# colour_long_edges(.data = distance_umap_s_curve, benchmark_value = benchmark_umap_s_curve,
#                   triangular_object = tr1_object_umap_s_curve, distance_col = distance)

trimesh_removed_umap_s_curve <- remove_long_edges(.data = distance_umap_s_curve, benchmark_value = benchmark_umap_s_curve,
                                     triangular_object = tr1_object_umap_s_curve, distance_col = distance)

trimesh_removed_umap_s_curve <- trimesh_removed_umap_s_curve +
  # xlab(expression(C[x]^{(2)})) + ylab(expression(C[y]^{(2)})) +
  # theme(axis.text = element_text(size = 5),
  #       axis.title = element_text(size = 7)) +
  geom_point(colour = "#33a02c", size = 0.5) +
  theme_linedraw() +
  theme(plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5), 
        axis.title.x = element_blank(), axis.title.y = element_blank(),
        axis.text.x = element_blank(), axis.ticks.x = element_blank(),
        axis.text.y = element_blank(), axis.ticks.y = element_blank(),
        panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
  annotate(geom = 'text', label = 'b', x = -Inf, y = Inf, hjust = -0.5, vjust = 1.5, size = 3)

# tour_umap_s_curve <- show_langevitour(df_all_umap_s_curve, df_bin_umap_s_curve, df_bin_centroids_umap_s_curve, benchmark_value = benchmark_umap_s_curve, distance = distance_umap_s_curve, distance_col = distance)

num_bins_trimap_s_curve <- 6
shape_val_trimap_s_curve <- calculate_effective_shape_value(.data = TriMAP_s_curve,
                                                            x = TriMAP1, y = TriMAP2) ## 1.259938
## To extract bin centroids
hexbin_data_object_trimap_s_curve <- extract_hexbin_centroids(nldr_df = TriMAP_s_curve, num_bins = num_bins_trimap_s_curve, shape_val = shape_val_trimap_s_curve, x = TriMAP1, y = TriMAP2)

df_bin_centroids_trimap_s_curve <- hexbin_data_object_trimap_s_curve$hexdf_data

TriMAP_data_with_hb_id_s_curve <- TriMAP_s_curve |>
  dplyr::mutate(hb_id = hexbin_data_object_trimap_s_curve$hb_data@cID)

## To generate a data set with high-D and 2D training data
df_all_trimap_s_curve <- dplyr::bind_cols(training_data |> dplyr::select(-ID), TriMAP_data_with_hb_id_s_curve)

## Averaged on high-D
df_bin_trimap_s_curve <- avg_highD_data(.data = df_all_trimap_s_curve)

## Triangulate bin centroids
tr1_object_trimap_s_curve <- triangulate_bin_centroids(df_bin_centroids_trimap_s_curve, x, y)
tr_from_to_df_trimap_s_curve <- generate_edge_info(triangular_object = tr1_object_trimap_s_curve)

# ggplot(df_bin_centroids_trimap_s_curve, aes(x = x, y = y)) +
#   geom_segment(data = tr_from_to_df_trimap_s_curve, aes(x = x_from, y = y_from, xend = x_to, yend = y_to)) +
#   geom_point(size = 2, colour = "#33a02c") +
#   coord_equal()

## Compute 2D distances
distance_trimap_s_curve <- cal_2D_dist(.data = tr_from_to_df_trimap_s_curve)

## To find the benchmark value
benchmark_trimap_s_curve <- find_benchmark_value(.data = distance_trimap_s_curve, distance_col = distance)

# colour_long_edges(.data = distance_trimap_s_curve, benchmark_value = benchmark_trimap_s_curve,
#                   triangular_object = tr1_object_trimap_s_curve, distance_col = distance)

trimesh_removed_trimap_s_curve <- remove_long_edges(.data = distance_trimap_s_curve, benchmark_value = benchmark_trimap_s_curve,
                                                    triangular_object = tr1_object_trimap_s_curve, distance_col = distance)

trimesh_removed_trimap_s_curve <- trimesh_removed_trimap_s_curve +
  # xlab(expression(C[x]^{(2)})) + ylab(expression(C[y]^{(2)})) +
  # theme(axis.text = element_text(size = 5),
  #       axis.title = element_text(size = 7)) +
  geom_point(colour = "#33a02c", size = 0.25) +
  theme_linedraw() +
  theme(plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5), 
        axis.title.x = element_blank(), axis.title.y = element_blank(),
        axis.text.x = element_blank(), axis.ticks.x = element_blank(),
        axis.text.y = element_blank(), axis.ticks.y = element_blank(),
        panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
  annotate(geom = 'text', label = 'd', x = Inf, y = Inf, hjust = 1.5, vjust = 1.5, size = 3)

# tour_trimap_s_curve <- show_langevitour(df_all_trimap_s_curve, df_bin_trimap_s_curve, df_bin_centroids_trimap_s_curve, benchmark_value = benchmark_trimap_s_curve, distance = distance_trimap_s_curve, distance_col = distance)

## PacMAP

num_bins_pacmap_s_curve <- 10
shape_val_pacmap_s_curve <- calculate_effective_shape_value(.data = PaCMAP_s_curve,
                                                            x = PaCMAP1, y = PaCMAP2) ## 1.259938
## To extract bin centroids
hexbin_data_object_pacmap_s_curve <- extract_hexbin_centroids(nldr_df = PaCMAP_s_curve, num_bins = num_bins_pacmap_s_curve, shape_val = shape_val_pacmap_s_curve, x = PaCMAP1, y = PaCMAP2)

df_bin_centroids_pacmap_s_curve <- hexbin_data_object_pacmap_s_curve$hexdf_data

PaCMAP_data_with_hb_id_s_curve <- PaCMAP_s_curve |>
  dplyr::mutate(hb_id = hexbin_data_object_pacmap_s_curve$hb_data@cID)

## To generate a data set with high-D and 2D training data
df_all_pacmap_s_curve <- dplyr::bind_cols(training_data |> dplyr::select(-ID), PaCMAP_data_with_hb_id_s_curve)

## Averaged on high-D
df_bin_pacmap_s_curve <- avg_highD_data(.data = df_all_pacmap_s_curve)

## Triangulate bin centroids
tr1_object_pacmap_s_curve <- triangulate_bin_centroids(df_bin_centroids_pacmap_s_curve, x, y)
tr_from_to_df_pacmap_s_curve <- generate_edge_info(triangular_object = tr1_object_pacmap_s_curve)

# ggplot(df_bin_centroids_pacmap_s_curve, aes(x = x, y = y)) +
#   geom_segment(data = tr_from_to_df_pacmap_s_curve, aes(x = x_from, y = y_from, xend = x_to, yend = y_to)) +
#   geom_point(size = 2, colour = "#33a02c") +
#   coord_equal()

## Compute 2D distances
distance_pacmap_s_curve <- cal_2D_dist(.data = tr_from_to_df_pacmap_s_curve)

## To find the benchmark value
benchmark_pacmap_s_curve <- find_benchmark_value(.data = distance_pacmap_s_curve, distance_col = distance)

# colour_long_edges(.data = distance_pacmap_s_curve, benchmark_value = benchmark_pacmap_s_curve,
#                   triangular_object = tr1_object_pacmap_s_curve, distance_col = distance)

trimesh_removed_pacmap_s_curve <- remove_long_edges(.data = distance_pacmap_s_curve, benchmark_value = benchmark_pacmap_s_curve,
                                     triangular_object = tr1_object_pacmap_s_curve, distance_col = distance)

trimesh_removed_pacmap_s_curve <- trimesh_removed_pacmap_s_curve +
  # xlab(expression(C[x]^{(2)})) + ylab(expression(C[y]^{(2)})) +
  # theme(axis.text = element_text(size = 5),
  #       axis.title = element_text(size = 7)) +
  geom_point(colour = "#33a02c", size = 0.5) +
  theme_linedraw() +
  theme(plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5), 
        axis.title.x = element_blank(), axis.title.y = element_blank(),
        axis.text.x = element_blank(), axis.ticks.x = element_blank(),
        axis.text.y = element_blank(), axis.ticks.y = element_blank(),
        panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
  annotate(geom = 'text', label = 'e', x = Inf, y = Inf, hjust = 1.5, vjust = 1.5, size = 3)

# tour_pacmap_s_curve <- show_langevitour(df_all_pacmap_s_curve, df_bin_pacmap_s_curve, df_bin_centroids_pacmap_s_curve, benchmark_value = benchmark_pacmap_s_curve, distance = distance_pacmap_s_curve, distance_col = distance)

## PHATE

num_bins_phate_s_curve <- 8
shape_val_phate_s_curve <- calculate_effective_shape_value(.data = PHATE_s_curve,
                                                           x = PHATE1, y = PHATE2) ## 1.259938
## To extract bin centroids
hexbin_data_object_phate_s_curve <- extract_hexbin_centroids(nldr_df = PHATE_s_curve, num_bins = num_bins_phate_s_curve, shape_val = shape_val_phate_s_curve, x = PHATE1, y = PHATE2)

df_bin_centroids_phate_s_curve <- hexbin_data_object_phate_s_curve$hexdf_data

## Identify bins with low-density
identify_rm_bins <- find_low_density_hexagons(df_bin_centroids_phate_s_curve, num_bins_phate_s_curve, benchmark_rm_hex = 0.06)

df_bin_centroids_phate_s_curve <- df_bin_centroids_phate_s_curve |>
  filter(!(hexID %in% identify_rm_bins))

PHATE_data_with_hb_id_s_curve <- PHATE_s_curve |>
  dplyr::mutate(hb_id = hexbin_data_object_phate_s_curve$hb_data@cID)

## To generate a data set with high-D and 2D training data
df_all_phate_s_curve <- dplyr::bind_cols(training_data |> dplyr::select(-ID), PHATE_data_with_hb_id_s_curve)

## Averaged on high-D
df_bin_phate_s_curve <- avg_highD_data(.data = df_all_phate_s_curve)

## Triangulate bin centroids
tr1_object_phate_s_curve <- triangulate_bin_centroids(df_bin_centroids_phate_s_curve, x, y)
tr_from_to_df_phate_s_curve <- generate_edge_info(triangular_object = tr1_object_phate_s_curve)

# ggplot(df_bin_centroids_phate_s_curve, aes(x = x, y = y)) +
#   geom_segment(data = tr_from_to_df_phate_s_curve, aes(x = x_from, y = y_from, xend = x_to, yend = y_to)) +
#   geom_point(size = 2, colour = "#33a02c") +
#   coord_equal()

## Compute 2D distances
distance_phate_s_curve <- cal_2D_dist(.data = tr_from_to_df_phate_s_curve)

## To find the benchmark value
benchmark_phate_s_curve <- find_benchmark_value(.data = distance_phate_s_curve, distance_col = distance)

# colour_long_edges(.data = distance_phate_s_curve, benchmark_value = benchmark_phate_s_curve,
#                   triangular_object = tr1_object_phate_s_curve, distance_col = distance)

trimesh_removed_phate_s_curve <- remove_long_edges(.data = distance_phate_s_curve, benchmark_value = benchmark_phate_s_curve,
                                                   triangular_object = tr1_object_phate_s_curve, distance_col = distance)

trimesh_removed_phate_s_curve <- trimesh_removed_phate_s_curve +
  # xlab(expression(C[x]^{(2)})) + ylab(expression(C[y]^{(2)})) +
  # theme(axis.text = element_text(size = 5),
  #       axis.title = element_text(size = 7)) +
  geom_point(colour = "#33a02c", size = 0.5) +
  theme_linedraw() +
  theme(plot.title = element_text(size = 7, hjust = 0.5, vjust = -0.5), 
        axis.title.x = element_blank(), axis.title.y = element_blank(),
        axis.text.x = element_blank(), axis.ticks.x = element_blank(),
        axis.text.y = element_blank(), axis.ticks.y = element_blank(),
        panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
  annotate(geom = 'text', label = 'c', x = -Inf, y = Inf, hjust = -0.5, vjust = 1.5, size = 3)

# tour_phate_s_curve <- show_langevitour(df_all_phate_s_curve, df_bin_phate_s_curve, df_bin_centroids_phate_s_curve, benchmark_value = benchmark_phate_s_curve, distance = distance_phate_s_curve, distance_col = distance)

```{r}
#| warning: false
#| echo: false
#| label: fig-modelScurve
#| fig-cap: Model.

trimesh_removed_tsne_s_curve + trimesh_removed_umap_s_curve + trimesh_removed_phate_s_curve + trimesh_removed_trimap_s_curve + trimesh_removed_pacmap_s_curve +
  #plot_annotation(tag_levels = 'a') +
  plot_layout(ncol = 5) &
  theme(legend.position='none', plot.tag = element_text(size = 8))
```

@@

The approach involves employing the K-nearest neighbors (KNN) algorithm to identify the nearest hexagonal bin centroid in the 2D space. Subsequently, the coordinates of this centroid are assigned as the low-dimensional predicted values for the test data in 2D space. It is noteworthy that traditional NLDR methods, such as tSNE, often lack a direct predict function, making our approach valuable for generating predicted values in the absence of such functionalities.

@@

### Prediction {#sec-prediction}

In this context, "prediction" denotes the 2D embedding generated for the NLDR technique.
The methodology encompasses identifying the nearest averaged high-D points in the high-D space for the test data, by computing high-D Euclidean distances. As the averaged high-D point corresponds to the lifting of the 2D model, determining its nearest counterpart allows us to map its hexagonal bin centroid coordinates. Consequently, these centroid coordinates serve as the assigned low-dimensional predicted values for the test data in the 2D space.

Some NLDR techniques, such as tSNE, often lack a direct prediction, making our approach valuable for generating predicted values in the absence of such functionalities.

### 2024-01-14 paper.tex
Add the prediction section where 2D embeddings are predicted.

\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[frame hidden, borderline west={3pt}{0pt}{shadecolor}, sharp corners, breakable, boxrule=0pt, interior hidden, enhanced]}{\end{tcolorbox}}\fi

@@

The outline of this paper is as follows. The
Section~\ref{sec-background} provides an detailed overview of dimension
reduction methods, triangulation, and tours. Building upon this
foundation, the Section~\ref{sec-methods} delves into the proposed
algorithm, \textbf{quollr}, and its implementation details. In
\textbf{?@sec-prediction}, discusses the effectiveness of the learned
low-dimensional manifold in accurately representing the complex
high-dimensional data. Following that, Section~\ref{sec-simpleex}
presents simple examples from simulations to illustrate the
functionality of the algorithm. Subsequently,
Section~\ref{sec-applications} showcases real-world applications of
\textbf{quollr} on different data sets, particularly in single-cell
RNA-seq data. These applications reveal insights into the performance
and trustworthiness of NLDR algorithms. We analyze the results to
identify situations where NLDR techniques may lead to misleading
interpretations. Finally, \textbf{?@sec-conclusions} concludes by
summarizing the findings and emphasizing the significance of the
proposed approach in tackling the challenges of high-dimensional data
visualization.

@@

@@

In this context, the term ``prediction values'' refers to the 2D
coordinates predicted for the NLDR technique. The approach involves
employing the K-nearest neighbors (KNN) algorithm to identify the
nearest hexagonal bin centroid in the 2D space. Subsequently, the
coordinates of this centroid are assigned as the low-dimensional
predicted values for the test data in 2D space. It is noteworthy that
traditional NLDR methods, such as t-SNE, often lack a direct predict
function, making our approach valuable for generating predicted values
in the absence of such functionalities.

@@

>>>>

\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[boxrule=0pt, interior hidden, breakable, enhanced, sharp corners, frame hidden, borderline west={3pt}{0pt}{shadecolor}]}{\end{tcolorbox}}\fi

@@

The outline of this paper is as follows. The
Section~\ref{sec-background} provides an detailed overview of dimension
reduction methods, triangulation, and tours. Building upon this
foundation, the Section~\ref{sec-methods} delves into the proposed
algorithm, \textbf{quollr}, and its implementation details. In
Section~\ref{sec-prediction}, discusses the effectiveness of the learned
low-dimensional manifold in accurately representing the complex
high-dimensional data. Following that, Section~\ref{sec-simpleex}
presents simple examples from simulations to illustrate the
functionality of the algorithm. Subsequently,
Section~\ref{sec-applications} showcases real-world applications of
\textbf{quollr} on different data sets, particularly in single-cell
RNA-seq data. These applications reveal insights into the performance
and trustworthiness of NLDR algorithms. We analyze the results to
identify situations where NLDR techniques may lead to misleading
interpretations. Finally, \textbf{?@sec-conclusions} concludes by
summarizing the findings and emphasizing the significance of the
proposed approach in tackling the challenges of high-dimensional data
visualization.

@@

{\centering \includegraphics{paper_files/figure-pdf/fig-modelScurve-1.pdf}

\caption{\label{fig-modelScurve}Model.}

@@

The approach involves employing the K-nearest neighbors (KNN) algorithm
to identify the nearest hexagonal bin centroid in the 2D space.
Subsequently, the coordinates of this centroid are assigned as the
low-dimensional predicted values for the test data in 2D space. It is
noteworthy that traditional NLDR methods, such as tSNE, often lack a
direct predict function, making our approach valuable for generating
predicted values in the absence of such functionalities.

@@

\hypertarget{sec-prediction}{%
\subsection{Prediction}\label{sec-prediction}}

In this context, ``prediction'' denotes the 2D embedding generated for
the NLDR technique. The methodology encompasses identifying the nearest
averaged high-D points in the high-D space for the test data, by
computing high-D Euclidean distances. As the averaged high-D point
corresponds to the lifting of the 2D model, determining its nearest
counterpart allows us to map its hexagonal bin centroid coordinates.
Consequently, these centroid coordinates serve as the assigned
low-dimensional predicted values for the test data in the 2D space.

Some NLDR techniques, such as tSNE, often lack a direct prediction,
making our approach valuable for generating predicted values in the
absence of such functionalities.
